### 2019-07-11 submission/manuscript.tex
Responding to Pat's comments and making additional edits

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether PPIs promote CDIs using a CDI mouse model.
Mice were treated with a high daily dose of the PPI omeprazole for 7
days prior to \emph{C. difficile} challenge and were compared to mice
that were treated with the antibiotic clindamycin or both omeprazole and
clindamycin. High-dose PPI treatment was maintained throughout the
experiment, which included \emph{C. difficile} challenge and the
following 9 days. We found that omeprazole was not sufficient to promote
\emph{C. difficile} colonization. When omeprazole treatment was combined
with the antibiotic, there was a mixed level of \emph{C. difficile}
colonization, where one cage of mice remained resistant and the other
cage was colonized. 16S rRNA sequencing analysis revealed that
omeprazole had minimal impact on the murine microbiota throughout the 16
days of PPI exposure. \emph{C. difficile} colonization of
clindamycin-treated mice was associated with changes in
\emph{Alistipes}, \emph{Barnesiella}, Porphyromonadaceae, and
Ruminococcaceae. These results suggest PPI treatment alone is not
sufficient to disrupt microbiota resistance to C. difficle infection in
mice that are normally resistant in the absence of antibiotic treatment.

@@

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but it is less clear what other factors contribute to
CDI incidence and recurrence. Interestingly, other medications besides
antibiotics have been shown to alter the microbiota. Proton pump
inhibitors (PPIs) are associated with alterations in the human
intestinal microbiota through observational and interventional studies
and PPI use has also been associated with CDI incidence and recurrence.
We evaluated the ability of a high dose of the PPI omeprazole to alter
the murine intestinal microbiota and disrupt colonization resistance to
\emph{C. difficile}. We found daily PPI treatment had minimal impact on
the murine intestinal microbiota and was not enough to promote \emph{C.
difficile} colonization after challenge. Further studies are needed to
determine whether other factors such as the composition of the starting
bacterial community, comorbidities, and use of additional medications
contribute to the association between PPIs and CDIs.

@@

Proton pump inhibitors (PPIs) are one of the most widely used drugs and
are available over the counter (1). PPIs are approved to treat a variety
of conditions including gastroesophageal reflux disease, ulcers, or as
prophylaxis for nonsteroidal anti-inflammatory drug (NSAID)-associated
gastrointestinal bleeding (2). Unfortunately, there have been a number
of adverse events associated with their use including small intestinal
bacterial overgrowth, decreased calcium absorption, and increased risk
of enteric infections (2). There's also been a number of large cohort
studies as well as interventional clinical trials that have demonstrated
specific alterations in the intestinal microbiome were associated with
PPI use (3, 4). PPI-associated microbiota changes have been attributed
to the ability of PPIs to increase stomach acid pH which may promote the
survival of oral and pathogenic bacteria (3, 4). Human fecal microbiota
changes with PPI use include increases in Enterococcaceae,
Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae
with decreases in Ruminococcaceae (3). An \emph{in vitro screen},
demonstrated that 3 PPIs inhibit the growth of 7-8 strains of
Lachnospiraceae and Ruminococcaceae, which represented
\textasciitilde{}16\% of the 40 intestinal bacteria strains that were
screened (5).\\
Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (6). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between PPI use and incidence or recurrence of CDIs (2,
4, 7, 8). However, most of the studies suggesting a link between PPIs
and \emph{C. difficile} are retrospective and did not evaluate the
microbiome (2, 7, 8). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use play a role in the
association between PPIs and CDI incidence or recurrence. Additionally,
epidemiological studies have a limited capacity to address potential
confounders and comorbidities in patients that were on PPIs and
developed CDIs or recurrent CDIs (7, 8).\\
Although there have been a few \emph{C. difficile} mouse model studies
that have demonstrated PPIs have some effect on CDIs with or without
additional antibiotic treatment, none of these studies looked at the
interplay with the intestinal microbiome (9--11). Here, we evaluated the
impact of a daily high dose PPI treatment on the murine microbiome and
susceptibility to \emph{C. difficile} colonization in relation to
clindamycin, an antibiotic that perturbs the microbiome enough to allow
\emph{C. difficile} to colonize but is mild enough that \emph{C.
difficile} is cleared within 10 days (12).

\subsection{Results}\label{results}

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test if PPI treatment alters the microbiome and promotes
susceptibility to CDIs, we gavaged mice daily with a high dose of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A)
and compared the PPI-treated mice to a group that received the
antibiotic, clindamycin, and a group that received both PPI and
clindamycin treatment. A principle coordinates analysis (PCoA) of the
Bray-Curtis distances over the initial 7 days of treatment revealed the
bacterial communities of PPI-treated mice remained relatively unchanged
(Figure 1B). Enterococcaceae, Lactobacillaceae, Micrococcaceae,
Staphylococcaceae, Streptococcaceae, and Ruminococcaceae are all familes
that have previously been impacted by PPIs in human studies, but we see
no significant fluctuations in the stool of our PPI-treated mice
throughout the course of the 16-day experiment (Figure 1C, Figure S1).
In contrast, the microbiomes from the 2 groups of mice that received
clindamycin started to shift away from the PPI-treated mice and the
samples from previous days one day after treatment (Figure 1D).

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} After 7 days of PPI treatment or 1 day
after clindamycin treatment, mice were challenged with \emph{C.
difficile} 630 spores. While all 4 of the clindamycin-treated mice were
colonized with \emph{C. difficile}, all of the PPI-treated mice were
resistant to \emph{C. difficile} colonization (Figure 2A).
Interestingly, only 1 cage of the PPI and clindamycin-treated mice were
colonized, while the other cage was not (Figure 2A). PCoAs of the
bacterial communities from the 3 groups of mice after \emph{C.
difficile} challenge, showed that the greatest shifts in bacterial
communities occurred in the clindamycin-treated mice (Figure 2B). Within
5 days, all of the mice had cleared \emph{C. difficile}, suggesting
there was no difference in rate of clearance for the mice that were
initially colonized with \emph{C. difficile} despite continuing to
receive PPI treatment throughout the course of the experiment. Our
results suggest that PPI treatment alone had minor effects on microbial
community resistance to \emph{C. difficile} colonization in mice.
Instead most of the differences between our 3 treatment groups appear to
be driven by clindamycin administration and included decreased
\emph{Alistipes}, \emph{Barnesiella}, Porhyromonadaceae,
Ruminococcaceae, taxa previously found to be altered in
clindamycin-treated mice that were challenged with \emph{C. difficile}
(6).

@@

Our results demonstrate that daily treatment with a high dose of the PPI
omeprazole had minimal impact on the stool microbiomes of mice and did
not promote susceptibility to \emph{C. difficile} colonization. In
contrast, clindamycin-treated mice were initially colonized, but cleared
the infection within 10 days. Combining PPI and clindamycin treatment
had a mixed effect, one cage was colonized with \emph{C. difficile} and
cleared within 10 days, while the other cage remained resistant.\\

@@

microbiome are comparable to another PPI mouse study that indicated PPIs
have more of an effect on the small intestinal microbiota compared to
the fecal microbiota (13). We did not find significant changes for the
taxa observed to be significantly impacted by PPI use in human studies.
However, 3 of the families that typically increase were absent or at low
abundance in the mice from the beginning and may be one contributing
factor to why we did not see any changes. Additionally, some of the
significance of PPI associations in human interventional trials appears
to be driven by a handful of specific taxa with overall differences on a
PCoA plot difficult to distinguish (14). Also, most of the studies to
date compare different individuals (PPI users versus non-users) or the
same individuals before and after treatment (3, 4), unfortunately these
limited microbiota snapshots may be skewed by day-to-day microbiota
variations (16) that would be revealed with more frequent longitudinal
sampling.\\
Previous work that administered 0.5 mg/kg of the PPI lansoprazole daily
for 2 weeks to mice and then challenged with \emph{C. difficile}
demonstrated that PPI treatment alone resulted in detectable \emph{C.
difficile} in the stool 1 week after challenge, but also showed there
was detectable \emph{C. difficile} in mice not treated with antibiotics
(10, 11). The presence of \emph{C. difficile} in mice that were not
treated with either antibiotics or PPIs could be attributed to the
higher dose of vegetative cells (10\textsuperscript{8} CFU) used to
challenge the mice and may partly explain their observation of \emph{C.
difficile} in PPI-treated mice (10, 11). We have previously shown mice
from our colony that are not given antibiotics are resistant to \emph{C.
difficile} 630 when challenged with 10\textsuperscript{3} spores (17).
Another group also used a higher dose of 10\textsuperscript{6} spores or
10\textsuperscript{7} vegetative cells to challenge antibiotic-treated
mice or mice treated with both the PPI esomeprazole (40mg/kg dose) for 2
days and antibiotics and demonstrated the antibiotic/PPI-treated mice
developed more severe CDIs (9).\\

@@

between PPIs and CDI incidence or recurrence. NSAID treatment can
increase CDI severity in mice (18). Given that PPIs are sometimes used
in combination with NSAIDs (2), it may be worth examining both
medications together using a mouse model. This study only looked at the
impact of PPIs with or without antibiotics, but future work may want to
incorporate age, other comorbidities and bacterial strains that are less
common in mice to evaluate whether PPIs in conjunction with other
factors can increase the risk of CDIs or recurrent CDIs.

@@

\textbf{Animals} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. 2-3 mice were housed per cage
and male and female mice were housed separately. All animal experiments
were approved by the University of Michigan Animal Care and Use
Committee (IACUC) under protocol number PRO00006983.

@@

\hypertarget{refs}{}
\hypertarget{ref-Singh2018}{}
1. \textbf{Singh A}, \textbf{Cresci GA}, \textbf{Kirby DF}. 2018. Proton
pump inhibitors: Risks and rewards and emerging consequences to the gut
microbiome. Nutrition in Clinical Practice \textbf{33}:614--624.
doi:\href{https://doi.org/10.1002/ncp.10181}{10.1002/ncp.10181}.

\hypertarget{ref-nehra2018proton}{}
2. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns, pp. 240--246. \emph{In} Mayo clinic proceedings. Elsevier.

\hypertarget{ref-Imhann2017}{}
3. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

@@

\hypertarget{ref-maier2018extensive}{}
5. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Schubert2015}{}
6. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against clostridium
difficile. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\hypertarget{ref-tariq2017association}{}
7. \textbf{Tariq R}, \textbf{Singh S}, \textbf{Gupta A}, \textbf{Pardi
DS}, \textbf{Khanna S}. 2017. Association of gastric acid suppression
with recurrent clostridium difficile infection: A systematic review and
meta-analysis. JAMA internal medicine \textbf{177}:784--791.

\hypertarget{ref-Elias2019}{}
8. \textbf{Elias E}, \textbf{Targownik LE}. 2019. The clinician's guide
to proton pump inhibitor related adverse events. Drugs
\textbf{79}:715--731.
doi:\href{https://doi.org/10.1007/s40265-019-01110-3}{10.1007/s40265-019-01110-3}.

\hypertarget{ref-hung2015proton}{}
9. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Kaur2007}{}
10. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
11. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-Jenior2018}{}
12. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
13. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
14. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
15. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-LlornsRico2019}{}
16. \textbf{Lloréns-Rico V}, \textbf{Raes J}. 2019. Tracking humans and
microbes. Nature \textbf{569}:632--633.
doi:\href{https://doi.org/10.1038/d41586-019-01591-y}{10.1038/d41586-019-01591-y}.

>>>>

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether PPIs promote CDIs using a CDI mouse model.
Mice were treated with a high daily dose of the PPI omeprazole for 7
days prior to \emph{C. difficile} challenge and were compared to mice
that were treated with the antibiotic clindamycin or both omeprazole and
clindamycin. High-dose PPI treatment was maintained throughout the
experiment, which included pre-treatment, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that omeprazole was not
sufficient to promote \emph{C. difficile} colonization. When omeprazole
treatment was combined with the antibiotic,one cage of mice remained
resistant to \emph{C. difficile} colonization, while the other cage was
colonized. 16S rRNA sequencing analysis revealed that omeprazole had
minimal impact on the murine microbiota throughout the 16 days of PPI
exposure. These results suggest PPI treatment alone is not sufficient to
disrupt microbiota resistance to \emph{C. difficle} infection in mice
that are normally resistant in the absence of antibiotic treatment.

@@

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but it is less clear what other factors contribute to
CDI incidence and recurrence. Interestingly, other medications besides
antibiotics have been shown to alter the microbiota. Proton pump
inhibitors (PPIs) are associated with alterations in the human
intestinal microbiota through observational and interventional studies
and PPI use has also been associated with CDI incidence and recurrence
in hospital and epidemiological studies. We evaluated the ability of a
high dose of the PPI omeprazole to alter the murine intestinal
microbiota and disrupt colonization resistance to \emph{C. difficile}.
We found daily PPI treatment had minimal impact on the murine intestinal
microbiota and was not enough to promote \emph{C. difficile}
colonization after challenge. Further studies are needed to determine
whether other factors such as the composition of the starting bacterial
community, comorbidities, and use of additional medications contribute
to the association between PPIs and CDIs that is seen in humans.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). PPIs are one of the most widely used drugs in
the U.S. and are available over the counter (6). PPIs are approved to
treat a variety of conditions including gastroesophageal reflux disease,
ulcers, or as prophylaxis for nonsteroidal anti-inflammatory drug
(NSAID)-associated gastrointestinal bleeding (3). Unfortunately, there
have been a number of adverse events associated with their use including
small intestinal bacterial overgrowth, decreased calcium absorption, and
increased risk of enteric infections (3). There have also been a number
of large cohort studies as well as interventional clinical trials that
have demonstrated specific alterations in the intestinal microbiome were
associated with PPI use (4, 7). PPI-associated microbiota changes have
been attributed to the ability of PPIs to increase stomach acid pH which
may promote the survival of oral and pathogenic bacteria (4, 7). Human
fecal microbiota changes with PPI use include increases in
Enterococcaceae, Lactobacillaceae, Micrococcaceae, Staphylococcaceae and
Streptococcaceae and decreases in Ruminococcaceae (7). An \emph{in vitro
screen} demonstrated that 3 PPIs inhibit the growth of 7-8 strains of
Lachnospiraceae and Ruminococcaceae, which represented
\textasciitilde{}16\% of the 40 intestinal bacteria strains that were
screened (8), suggesting direct action on bacteria might also play a
role.\\
However, most of the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate the microbiome (2, 3,
5). Thus, it is unclear whether the gastrointestinal microbiome changes
associated with PPI use play a role in the association between PPIs and
CDI incidence or recurrence. Additionally, epidemiological studies have
a limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of a daily high dose PPI treatment on the
murine microbiome and susceptibility to \emph{C. difficile} colonization
in relation to clindamycin, an antibiotic that perturbs the microbiome
enough to allow \emph{C. difficile} to colonize but is mild enough that
\emph{C. difficile} is cleared within 10 days (9).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test if PPI treatment alters the microbiome and promotes
susceptibility to CDIs, we gavaged mice daily with a high dose of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
PPI-treated mice were compared to a group of mice that received the
antibiotic, clindamycin, and a group that received both PPI and
clindamycin treatment. A principle coordinates analysis (PCoA) of the
Bray-Curtis distances over the initial 7 days of treatment revealed the
bacterial communities of PPI-treated mice remained relatively unchanged
(Figure 1B). Enterococcaceae, Lactobacillaceae, Micrococcaceae,
Staphylococcaceae, Streptococcaceae, and Ruminococcaceae are all familes
that have previously been impacted by PPIs in human studies, but we
observed no significant fluctuations in the stool of our PPI-treated
mice throughout the course of the 16-day experiment (Figure 1C, Figure
S1). In contrast, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that high dose PPI treatment alone had no
significant impact on the murine fecal bacterial community.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether high dose PPI
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After 7 days of PPI treatment or 1 day after clindamycin
treatment, mice were challenged with \emph{C. difficile} 630 spores.
While all 4 of the clindamycin-treated mice were colonized with \emph{C.
difficile}, all of the PPI-treated mice were resistant to \emph{C.
difficile} colonization (Figure 2A). Interestingly, only 1 cage of the
PPI and clindamycin-treated mice were colonized, while the other cage
was not (Figure 2A). PCoAs of the bacterial communities from the 3
groups of mice after \emph{C. difficile} challenge, showed that the
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had
cleared \emph{C. difficile}, suggesting there was no difference in rate
of clearance for the mice that were initially colonized with \emph{C.
difficile} despite continuing to receive PPI treatment throughout the
course of the experiment. Our results suggest that PPI treatment alone
had minor effects on microbial community resistance to \emph{C.
difficile} colonization in mice. Instead most of the differences between
our 3 treatment groups appear to be driven by clindamycin administration
and included decreased \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered
in clindamycin-treated mice that were challenged with \emph{C.
difficile} (1). These findings demonstrated that high dose PPI treatment
did not promote susceptibility to \emph{C. difficile} colonization.

@@

@@

microbiome were comparable to another PPI mouse study that indicated
PPIs have more of an effect on the small intestinal microbiota compared
to the fecal microbiota (10). We did not find significant changes for
the taxa observed to be significantly impacted by PPI use in human
studies. However, 3 of the families that typically increase were absent
or at low abundance in the mice from the beginning and may be one
contributing factor to why we did not see any changes. Additionally,
some of the significance of PPI associations in human interventional
trials appears to be driven by a handful of specific taxa with overall
differences on a PCoA plot difficult to distinguish (11). Also, most of
the studies to date compareed different individuals (PPI users versus
non-users) or the same individuals before and after treatment (4, 7),
unfortunately these limited microbiota snapshots may be skewed by
day-to-day microbiota variations (13) that would be revealed with more
frequent longitudinal sampling.\\
ALthough there have been a few \emph{C. difficile} mouse model studies
that have demonstrated PPIs have some effect on CDIs with or without
additional antibiotic treatment, none of these studies looked at the
interplay with the intestinal microbiome (14--16). Additionally, there
were some key methodical differences between the previous studies and
our own. Previous work that administered 0.5 mg/kg of the PPI
lansoprazole daily for 2 weeks to mice and then challenged with \emph{C.
difficile} demonstrated that PPI treatment alone resulted in detectable
\emph{C. difficile} in the stool 1 week after challenge, but also showed
there was detectable \emph{C. difficile} in mice not treated with
antibiotics (14, 15). The presence of \emph{C. difficile} in mice that
were not treated with either antibiotics or PPIs could be attributed to
the higher dose of vegetative cells (10\textsuperscript{8} CFU) used to
challenge the mice and may partly explain the observation of \emph{C.
difficile} in PPI-treated mice (14, 15). We have previously shown mice
from our colony that were not given antibiotics were resistant to
\emph{C. difficile} 630 when challenged with 10\textsuperscript{3}
spores (17). Another group also used a higher dose of
10\textsuperscript{6} spores or 10\textsuperscript{7} vegetative cells
to challenge antibiotic-treated mice or mice treated with both the PPI
esomeprazole (40mg/kg dose) for 2 days and antibiotics and demonstrated
the antibiotic/PPI-treated mice developed more severe CDIs (16).\\

@@

between PPIs and CDI incidence or recurrence. For example, NSAID
treatment can increase CDI severity in mice (18) and PPIs are sometimes
used in combination with NSAIDs (3). This study addressed the impact of
PPIs with or without antibiotics on a murine model of CDI, and found
PPIs did not promote \emph{C. difficile} colonization. Future studies
are needed to determine whether age, other comorbidities and bacterial
strains that are less common in mice can increase the risk of CDIs or
recurrent CDIs when combined with PPI treatment.

@@

\textbf{Animals} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was spit between 2 cages with 2-3 mice housed per cage and male and
female mice were housed separately. All animal experiments were approved
by the University of Michigan Animal Care and Use Committee (IACUC)
under protocol number PRO00006983.

@@

\hypertarget{refs}{}
\hypertarget{ref-Schubert2015}{}
1. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against clostridium
difficile. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\hypertarget{ref-tariq2017association}{}
2. \textbf{Tariq R}, \textbf{Singh S}, \textbf{Gupta A}, \textbf{Pardi
DS}, \textbf{Khanna S}. 2017. Association of gastric acid suppression
with recurrent clostridium difficile infection: A systematic review and
meta-analysis. JAMA internal medicine \textbf{177}:784--791.

\hypertarget{ref-nehra2018proton}{}
3. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns, pp. 240--246. \emph{In} Mayo clinic proceedings. Elsevier.

@@

\hypertarget{ref-Elias2019}{}
5. \textbf{Elias E}, \textbf{Targownik LE}. 2019. The clinician's guide
to proton pump inhibitor related adverse events. Drugs
\textbf{79}:715--731.
doi:\href{https://doi.org/10.1007/s40265-019-01110-3}{10.1007/s40265-019-01110-3}.

\hypertarget{ref-Singh2018}{}
6. \textbf{Singh A}, \textbf{Cresci GA}, \textbf{Kirby DF}. 2018. Proton
pump inhibitors: Risks and rewards and emerging consequences to the gut
microbiome. Nutrition in Clinical Practice \textbf{33}:614--624.
doi:\href{https://doi.org/10.1002/ncp.10181}{10.1002/ncp.10181}.

\hypertarget{ref-Imhann2017}{}
7. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

\hypertarget{ref-maier2018extensive}{}
8. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Jenior2018}{}
9. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
10. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
11. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
12. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-LlornsRico2019}{}
13. \textbf{Lloréns-Rico V}, \textbf{Raes J}. 2019. Tracking humans and
microbes. Nature \textbf{569}:632--633.
doi:\href{https://doi.org/10.1038/d41586-019-01591-y}{10.1038/d41586-019-01591-y}.

\hypertarget{ref-Kaur2007}{}
14. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
15. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
16. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.
